---
title: "Data Wrangling Steps"
author: "Anthony Stachowitz"
date: "July 11, 2018"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(readxl)
```

### 1. Load dataset

This loads the individual datasets.   
```{r}
income_data <- read_excel("C:/Users/Anthony/Desktop/R studio projects/income/income_data.xlsx", col_names = FALSE)
```

### 2. Add headers to the data set.

The data set does not have headers. I will add headers durring this step.
```{r}
names(income_data) <- c("Age", "WorkClass", "FnlWgt", "Education", "Education-num", "MaritalStatus", "Occupation", "Relationship", "Race", "Sex", "CapitalGain", "CapitalLoss", "HoursPerWeek", "NativeCountry", "Salary")
```

### 3. Add column to convert less than $50,000 to 1 and more than $50,000 to 0.


```{r}
income_data <- mutate(income_data, LessThen_50 = ifelse(grepl(">50K", Salary), 0, 1))
```

### 4. Convert characters to factors

Converted WorkClass, Education, MaritalStatus, and Occupation

```{r}
income_data$Occupation <- as.factor(income_data$Occupation)
income_data$WorkClass <- as.factor(income_data$WorkClass)
income_data$Education <- as.factor(income_data$Education)
income_data$MaritalStatus <- as.factor(income_data$MaritalStatus)
```

### 5. Remove the data that is not being used.

Remove columns to make the easier to work with.

```{r}
income_data <- income_data[, -c(1, 3, 5, 8:14)]
income_data
```

### 6. Check for NA's in the data

```{r}
sapply(income_data,function(x) sum(is.na(x)))
```

