---
title: "Data Wrangling Steps"
author: "Anthony Stachowitz"
date: "July 11, 2018"
output:
  html_document: default
  pdf_document: default
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(readxl)
```

### 1. Load dataset

This loads the individual datasets.   
```{r}
income_data <- read_excel("C:/Users/Anthony/Desktop/R studio projects/income/income_data.xlsx", col_names = FALSE)
```

### 2. Add headers to the data set.

The data set does not have headers. I will add headers durring this step.
```{r}
names(income_data) <- c("Age", "WorkClass", "FnlWgt", "Education", "Education-num", "MaritalStatus", "Occupation", "Relationship", "Race", "Sex", "CapitalGain", "CapitalLoss", "HoursPerWeek", "NativeCountry", "Salary")
```

### 3. Add column to convert less than $50,000 to 1 and more than $50,000 to 0.


```{r}
income_data <- mutate(income_data, LessThen_50 = ifelse(grepl(">50K", Salary), 0, 1))
```

### 4. Convert characters to factors

Converted WorkClass, Education, MaritalStatus, and Occupation

```{r}
income_data$Occupation <- as.factor(income_data$Occupation)
income_data$WorkClass <- as.factor(income_data$WorkClass)
income_data$Education <- as.factor(income_data$Education)
income_data$MaritalStatus <- as.factor(income_data$MaritalStatus)
```

### 5. Remove the data that is not being used.

Remove columns to make the easier to work with.

```{r}
income_data <- income_data[, -c(1, 3, 5, 8:15)]
income_data
```



### 6. turn missing values into NA's

```{r}
income_data[ income_data == "?" ] <- NA
```

### 7. Re-check for NA's

```{r}
sapply(income_data,function(x) sum(is.na(x)))
```

### 8. Remove rows of data that had NA's

```{r}
income_data <- income_data[complete.cases(income_data), ]
```

### 9. Recheck and data structure

```{r}
sapply(income_data,function(x) sum(is.na(x)))
str(income_data)
```




```{r}
library("RColorBrewer")
library("ggplot2")

ed_graph <- income_data %>% 
  group_by(Education) %>% 
  summarize(education_count = n()) %>% 
  arrange(desc(education_count))

ed_graph$Education <- factor(ed_graph$Education, levels = ed_graph$Education[order(ed_graph$Education)])
colourCount = length(unique(ed_graph$Education))
fill_various <- colorRampPalette(brewer.pal(8, "Dark2"))

ed_graph %>% 
  filter(Education != "NA") %>% 
  ggplot(aes(x = Education, y = education_count, fill = Education)) + 
  geom_bar(stat = "identity") + 
  coord_flip() +
  geom_text(aes(label = education_count), size = 3, color = "black", hjust = -0.1) + 
  labs(x = "Education Level", y = "Number of People", title = "Education Level sample of the Census") + 
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5)) + 
  ylim(0, max(ed_graph$education_count + 100)) + 
  scale_fill_manual(values = fill_various(colourCount))

```



```{r}

work_graph <- income_data %>% 
  group_by(WorkClass) %>% 
  summarize(work_count = n()) %>% 
  arrange(desc(work_count))

work_graph$WorkClass <- factor(work_graph$WorkClass, levels = work_graph$WorkClass[order(work_graph$WorkClass)])
colourCount = length(unique(work_graph$WorkClass))
fill_various <- colorRampPalette(brewer.pal(8, "Dark2"))

work_graph %>% 
  filter(WorkClass != "NA") %>% 
  ggplot(aes(x = WorkClass, y = work_count, fill = WorkClass, width = .4)) + 
  geom_bar(stat = "identity") + 
  geom_text(aes(label = work_count), size = 3, color = "black", hjust = 0.5, vjust = -0.7) + 
  labs(x = "Class of Work", y = "Number of People", title = "Class of Work Level sample of the Census") + 
  theme(legend.position = "none", plot.title = element_text(hjust = .5)) + 
  ylim(0, max(work_graph$work_count + 100)) + 
  scale_fill_manual(values = fill_various(colourCount))



```



```{r}

marital_graph <- income_data %>% 
  group_by(MaritalStatus) %>% 
  summarize(marital_count = n()) %>% 
  arrange(desc(marital_count))

marital_graph$MaritalStatus <- factor(marital_graph$MaritalStatus, levels = marital_graph$MaritalStatus[order(marital_graph$MaritalStatus)])
colourCount = length(unique(marital_graph$MaritalStatus))
fill_various <- colorRampPalette(brewer.pal(8, "Dark2"))

marital_graph %>% 
  filter(MaritalStatus != "NA") %>% 
  ggplot(aes(x = MaritalStatus, y = marital_count, fill = MaritalStatus, width = .4)) + 
  geom_bar(stat = "identity") +
  coord_flip() +
  geom_text(aes(label = marital_count), size = 3, color = "black", hjust = 0, vjust = -1.8) + 
  labs(x = "Marital Status", y = "Number of People", title = "Marital Status Level sample of the Census") + 
  theme(legend.position = "none", plot.title = element_text(hjust = .5)) + 
  ylim(0, max(marital_graph$marital_count + 100)) + 
  scale_fill_manual(values = fill_various(colourCount))


```



```{r}

Occup_graph <- income_data %>% 
  group_by(Occupation) %>% 
  summarize(Occupation_count = n()) %>% 
  arrange(desc(Occupation_count))

Occup_graph$Occupation <- factor(Occup_graph$Occupation, levels = Occup_graph$Occupation[order(Occup_graph$Occupation)])
colourCount = length(unique(Occup_graph$Occupation))
fill_various <- colorRampPalette(brewer.pal(8, "Dark2"))

Occup_graph %>% 
  filter(Occupation != "NA") %>% 
  ggplot(aes(x = Occupation, y = Occupation_count, fill = Occupation)) + 
  geom_bar(stat = "identity") + 
  coord_flip() +
  geom_text(aes(label = Occupation_count), size = 3, color = "black", hjust = -0.1) + 
  labs(x = "Occupation Level", y = "Number of People", title = "Occupation Level sample of the Census") + 
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5)) + 
  ylim(0, max(Occup_graph$Occupation_count + 100)) + 
  scale_fill_manual(values = fill_various(colourCount))



```